---
title: "OPTIMUM-Neuro conference call 2020-04-03"
output: html_document
always_allow_html: yes
---

```{r setup, include = FALSE, warning = FALSE}

# load graphics libraries
    library(plyr)
    library(ggplot2)
    library(reshape2)
    library(tidyverse) 
    library(knitr)
    library(kableExtra)
    library(plotly)
    library(stringr)
    library(bsselectR)
    library(lubridate)
    library(textclean)


```

```{r data cleaning, include = FALSE, warning = FALSE}

source('functions.R')
#read in data
  mri <- read.csv('/scratch/gherman/OPTIMUMMainDatabaseF_DATA_2020-04-03_0955.csv',na.strings=c("","NA")) #contains all but MRI
  df <- read.csv('/scratch/gherman/OPTIMUMMainDatabaseF_DATA_2020-04-03_0959.csv', na.strings=c("","NA")) #contains just MRI
  targets <- read.csv('info/OPT_recruitTargets.csv')
  #make sure target data is in proper format
  targets$month <- as.character(targets$month)
  
    sites <- factor(x = c('CU', 'LA', 'UP', 'UT', 'WU'), levels = c('CU', 'LA', 'UP', 'UT', 'WU'))
  

  
  df <- prepare_init_df(df, mri, sites)
  
```
>We are currently in month `r targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'month_count']` of the study.  

This report is meant to provide a quick overview of the OPTIMUM-Neuro study, to facilitate discussion on its bi-weekly conference calls. All data summarized here derives from the 'tracking' arm of the WashU REDCap database (arm 6), and thus is only as accurate as the data captured there. This report was generated on `r as.character(Sys.time())`
<br> 

------
<br> 




```{r recruitment, echo = FALSE, include = FALSE, warning = FALSE}

#make a smaller dataframe
recruit_df <- df[, c(
  "record_id", 
  "site",
  "timepoint",
  
  "meta_consent_date", 
  "meta_terminate_date",
  "meta_terminate_reason",
  
  "plasma_blood_status", 
  "plasma_blood_date",
  
  "mr_t1",
  "mr_date",
  
  "dkefs_complete",
  "dkefs_date_admin",
  "rbans_complete",
  "rbans_date_admin")] 

recruit_df <- prepare_recruit_df(recruit_df)
 
 
```
 
```{r enrolment, echo = FALSE, include = FALSE, warning = FALSE}
#add column that indicates if enrolled (where enrollment is 2 of 3 mri, blood, neuropsych)
  enroll_df <- prepare_enroll_df(recruit_df) #make a new dataframe (did this because want new DF to consider only baseline)

 
 
#merge back dataframes #what's the purpose of merging it here?
recruit_df <- merge(recruit_df, enroll_df, by = c('record_id', 'timepoint'), all.x = TRUE)
  
```

```{r, echo = FALSE, include = FALSE, warning = FALSE}


#note to self.... the enrollmrimth thing doesnt really work as the only indicator, it should really be combined with mr_t1 (which is enroll_mri i guess) to give an accurate picture of numbers. 


fu_due_list <- recruit_df%>% select(record_id,mri_fu_due, np_fu_due, bld_fu_due, mri_fu_7_mos, np_fu_7_mos,bld_fu_7_mos, enroll_mri_init,mri_fu_mth, mri_fu_mthT ) %>% filter(!is.na(mri_fu_due) | !is.na( np_fu_due) | !is.na( bld_fu_due)| !is.na(mri_fu_7_mos)| !is.na(np_fu_7_mos) | !is.na(bld_fu_7_mos))

fu_due_list <- fu_due_list[!duplicated(fu_due_list$record_id),]

#maybe I actually need to compare this to how many ppl actuvally received their init mri
fu_df <- recruit_df %>% filter(.,timepoint=="6 mth FU") %>% select(.,-c(enroll_mri, enroll_mri_date)) 

fu_df <- prep_fu_df(fu_df) %>% merge(.,fu_due_list, by=c('record_id'))
  
```


###Recruitment


```{r recruit_table, warning = FALSE, echo = FALSE}

recruit_table <- make_recruit_table(recruit_df, sites, targets)

#turns NAs into dashes
options(knitr.kable.NA = '--')

#table
kable(recruit_table, "html", escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", "condensed", full_width = T) %>%
  add_header_above(c(" ", "CU" = 2, "LA" = 2, "UP" = 2, "UT" = 2, "WU" = 2, " ")) %>%
  group_rows('Consent[note]', 1, 2, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Enrollment[note]", 3, 4, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Completion[note]", 5, 6, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Termination[note]", 7, 9, label_row_css = "background-color: #ddd; color: 999;") %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  add_footnote(c("`Consent` counts all participants who signed an OPTIMUM-Neuro consent form.", 
                 "`Enrollment` counts all participants who completed ≥2/3 of the MRI, bloodwork, and neuropsych at the baseline visit.",
                 "`Completion` counts all enrolled participants who completed the 24 mth FU visit.", 
                 "`Termination` counts participants coded as: 'not eligible for randomization', 'withdrew', 'lost to follow-up', 'death', 'investigator discretion', and 'other'. Termination does not capture participants who completed the study, i.e., this category is independent from `Completion`."), 
               notation = "number") %>%
  footnote(general = "Current month captures calendar month. The `%` columns calculate percentage of target, with all sites having common targets.")

```

------

##Progress Reports {.tabset}
Please switch between the tabs for baseline and 6 month follow-up numbers.

<br>

###Enrollment breakdown

```{r enrollment_table, echo = FALSE, warning = FALSE}

  enroll_table <- make_enroll_table(recruit_df, sites, targets)

#add row names to make like demo table
enroll_table <- cbind(metric = rownames(enroll_table), enroll_table)
rownames(enroll_table) <- NULL

#turns NAs into dashes
options(knitr.kable.NA = '--')

#table
kable(enroll_table, "html", escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", "condensed", full_width = T) %>%
  add_header_above(c(" ", "CU" = 2, "LA" = 2, "UP" = 2, "UT" = 2, "WU" = 2, " ")) %>%
  group_rows('Neuropsych[note]', 1, 2, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Blood[note]", 3, 4, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("MRI[note]", 5, 6, label_row_css = "background-color: #ddd; color: 999;") %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  add_footnote(c(
     "`Neuropsych` counts all participants who completed both the RBANS and D-KEFS in full.",
     "`Blood` counts all participants who submitted one aliquot of plasma.",
     "`MRI` counts all participants who completed the T1."), 
               notation = "number") %>%
  footnote(general = "Current month captures calendar month. The `%` columns calculate percentage of target, with all sites having common targets.")
              
```

###6 Month Follow-up 


Here, the % column indicates retention - this is calculated as the percent of participants who received an evaluation at baseline who have also received their 6 month follow-up evaluation. 

```{r, echo = FALSE, warning = FALSE}

fu_table <- make_fu_table(fu_df, sites,targets)

#add row names to make like demo table
fu_table <- cbind(metric = rownames(fu_table), fu_table)
rownames(fu_table) <- NULL

#turns NAs into dashes
options(knitr.kable.NA = '--')

#table
kable(fu_table, "html", escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", "condensed", full_width = T) %>%
  add_header_above(c(" ", "CU" = 2, "LA" = 2, "UP" = 2, "UT" = 2, "WU" = 2, " ")) %>%
  group_rows('Neuropsych[note]', 1, 2, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Blood[note]", 3, 4, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("MRI[note]", 5, 6, label_row_css = "background-color: #ddd; color: 999;") %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  add_footnote(c(
     "`Neuropsych` counts all participants who completed both the follow-up RBANS and D-KEFS in full.",
     "`Blood` counts all participants who submitted one aliquot of plasma for bloodwork follow-up.",
     "`MRI` counts all participants who completed the T1 follow-up."), 
               notation = "number") %>%
  footnote(general = "Current month captures calendar month. The `%` columns describe percent retention - what proportion of participants who received their baseline scan have received their 6 month follow-up in time.")
```

##

------


###Demographics 

<br>

```{r demographics, echo = FALSE, warning = FALSE}

demo_table <- make_demo_table(df, sites)

```

```{r demographics_table, echo = FALSE, warning = FALSE}

demo_table$p <- as.numeric(demo_table$p)

#table
demo_table %>%
  mutate(
    metric = row.names(demo_table),
    p = cell_spec(p, "html", color = ifelse(p < 0.05, "red", "black"))
    ) %>%
  select(metric, CU, LA, UP, UT, WU, p) %>%
  kable('html', escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", full_width = T) %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  footnote(general = "These values represent all participants consented at the OPTIMUM-Neuro baseline visit. It does not take into account the participants who consented but did not meet enrollment criteria, and/or who later terminated. Values in brackets are standard deviation. The _p_ values result from Chi-squared for categorical variables and omnibus ANOVA for continuous variables.")

rm(demo_df, demo_table)

```




