---
title: "OPTIMUM-Neuro conference call 2019-02-15"
output: html_document
always_allow_html: yes
---

```{r setup, include = FALSE, warning = FALSE}

# load graphics libraries
    library(plyr)
    library(ggplot2)
    library(reshape2)
    library(tidyverse)
    library(knitr)
    library(kableExtra)
    library(plotly)
    library(stringr)
    library(bsselectR)
    library(lubridate)
    library(textclean)

    #setwd('/projects/gherman/OPT_callReport')

```

```{r data cleaning, include = FALSE, warning = FALSE}

#read in data
  df <- read.csv('/home/gherman/Downloads/OPTIMUMMainDatabaseF_DATA_2019-04-05_1314.csv', na.strings=c("","NA")) #contains all but MRI
  mri <- read.csv('/home/gherman/Downloads/OPTIMUMMainDatabaseF_DATA_2019-04-05_1313.csv', na.strings=c("","NA")) #contains just MRI
  targets <- read.csv('info/OPT_recruitTargets.csv')
  #make sure target data is in proper format
  targets$month <- as.character(targets$month)
  
  df <- prepare_init_df(df, mri)
  
```
>We are currently in month `r targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'month_count']` of the study.  

This report is meant to provide a quick overview of the OPTIMUM-Neuro study, to facilitate discussion on its bi-weekly conference calls. All data summarized here derives from the 'tracking' arm of the WashU REDCap database (arm 6), and thus is only as accurate as the data captured there.
<br> 

------
<br> 

###Recruitment

```{r recruitment, echo = FALSE, include = FALSE, warning = FALSE}

#make a smaller dataframe
recruit_df <- df[, c(
  "record_id", 
  "site",
  "timepoint",
  
  "meta_consent_date", 
  "meta_terminate_date",
  "meta_terminate_reason",
  
  "plasma_blood_status", 
  "plasma_blood_date",
  
  "mr_t1",
  "mr_date",
  
  "dkefs_complete",
  "dkefs_date_admin",
  "rbans_complete",
  "rbans_date_admin")] 

recruit_df <- prepare_recruit_df(recruit_df)
 
 
```
 
```{r enrolment, echo = FALSE, include = FALSE, warning = FALSE}
#add column that indicates if enrolled (where enrollment is 2 of 3 mri, blood, neuropsych)
  enroll_df <- prepare_enroll_df(recruit_df) #make a new dataframe (did this because want new DF to consider only baseline)

 
 
#merge back dataframes 
recruit_df <- merge(recruit_df, enroll_df, by = c('record_id', 'timepoint'), all.x = TRUE)
  
```



```{r attempt_to_track_6mth_fu , echo = FALSE, include = FALSE, warning = FALSE}


#gabi attempt IS HERE


#note to self.... the enrollmrimth thing doesnt really work as the only indicator, it should really be combined with mr_t1 (which is enroll_mri i guess) to give an accurate picture of numbers. 



fu_df <- recruit_df %>% filter(.,timepoint=="6 mth FU") %>% select(.,-c(enroll_mri, enroll_mri_date))

fu_df$currentYm <- Sys.Date() #add current sysdate to df
fu_df$currentYm_str <- as.character(substr(fu_df$currentYm, 1, 7)) #store month and year, as character


#first, check if T1 complete, and if date is this month
  #note - here, 'complete' means just T1 complete
  names(fu_df)[names(fu_df) == "mr_t1"] <- "enroll_mri" #change name of status column
  names(fu_df)[names(fu_df) == "mr_date"] <- "enroll_mri_date" #change name of date column
  
  
  
  fu_df$enroll_mri <- ifelse(fu_df$enroll_mri == 1, 1, NA) #turn no to NA
  fu_df$enroll_mri_date <- as.Date(fu_df$enroll_mri_date, format = "%Y-%m-%d", origin = "1970-01-01") #convert date to datetime format
  fu_df$enroll_mri_mth <- as.character(substr(fu_df$enroll_mri_date, 1, 7)) #store month and year, as character
  fu_df$enroll_mri_mthT <- fu_df$enroll_mri_mth %in% fu_df$currentYm_str #logical T or F- was MRI this month?  #OKAY SO they must also have T1 completed
  
  
  # if they got an MRI this month then: fu_df %>% filter(enroll_mri==1, enroll_mri_mthT==TRUE)

#second, check if blood complete, and if date is this month
  #note - here, 'complete' means that just the plasma is complete (don't care about serum)
  fu_df$enroll_bld <- ifelse(fu_df$plasma_blood_status.x == 1, 1, NA) #turn no and partial to NA
  fu_df$enroll_bld_date <- substr(fu_df$plasma_blood_date.x, 1, 11)#cut out the time contained in datetime
  fu_df$enroll_bld_date <- as.Date(fu_df$enroll_bld_date, format = "%Y-%m-%d", origin = "1970-01-01") #convert to datetime 
  fu_df$enroll_bld_mth <- as.character(substr(fu_df$enroll_bld_date, 1, 7)) #store month and year, as character
  fu_df$enroll_bld_mthT <- fu_df$enroll_bld_mth %in% fu_df$currentYm_str #logical T or F - was blood this month?    
  
#third, check if neuropsych complete, and if data is this month
  #note - here, 'complete' means rbans and dkefs done in entirity
  #make new columns with completion and date data for 2 assessments we care about - dkefs and rbans
  fu_df$dkefs <- ifelse(fu_df$dkefs_complete.x == 1, 1, NA) #turn no and partial to NA
  fu_df$enroll_dkefs_date <- as.Date(fu_df$dkefs_date_admin.x, format = "%Y-%m-%d", origin = "1970-01-01") #convert date to datetime format
  fu_df$enroll_dkefs_mth <- as.character(substr(fu_df$enroll_dkefs_date, 1, 7)) #store month and year, as character
  fu_df$enroll_dkefs_mthT <- fu_df$enroll_dkefs_mth %in% fu_df$currentYm_str #logical T or F- was dkefs this month?

  fu_df$rbans <- ifelse(fu_df$rbans_complete.x == 1, 1, NA) #turn no and partial to NA
  fu_df$enroll_rbans_date <- as.Date(fu_df$rbans_date_admin.x, format = "%Y-%m-%d", origin = "1970-01-01") #convert date to datetime format
  fu_df$enroll_rbans_mth <- as.character(substr(fu_df$enroll_rbans_date, 1, 7)) #store month and year, as character
  fu_df$enroll_rbans_mthT <- fu_df$enroll_rbans_mth %in% fu_df$currentYm_str #logical T or F- was rbans this month?
  
  #take the sum of columns - want both of the central neuropsych assessments (dkefs, rbans) to be complete
  fu_df[, c('dkefs', 'rbans')] <- sapply(fu_df[, c('dkefs', 'rbans')], as.numeric) #change to numeric
  fu_df$np_count <- rowSums(fu_df[,c('dkefs', 'rbans')]) #sum

  #make an enrolled column for neuropsych - if sum is 2 
  fu_df$enroll_np <- ifelse(fu_df$np_count == 2, 1, 0)
  
  #identify the second date in ordered dates - i.e., when enrollment for np criteria is met
  np_dates <- c('dkefs_date.x', 'rbans_date.x')
  fu_df$enroll_np_date <- NA
  for (n in 1:nrow(fu_df)){
    dates <- fu_df[n, np_dates] # pull out NP assessment dates
    dates <- dates[order(dates)] # order dates
      if (length(dates[!is.na(dates)]) == 2) { # if there are 2 assesments
        fu_df$enroll_np_date[n] <- dates[[2]]   # pull out the second assessment date
      } else {
        fu_df$enroll_np_date[n] <- NA         # otherwise print 'NA'
      }
    }

  #determine if np criteria was met this month
  fu_df$enroll_np_date <- as.Date(fu_df$enroll_np_date, origin = "1970-01-01") #turn back into date format
  fu_df$enroll_np_mth <- as.character(substr(fu_df$enroll_np_date, 1, 7)) #store month and year, as character
  fu_df$enroll_np_mthT <- fu_df$enroll_np_mth %in% fu_df$currentYm_str #logical T or F- was criteria met this month?

  #determine if enrollment criteria is met, i.e., at least 2/3 of mri, blood, and np completed
  fu_df$enroll <- rowSums(fu_df[,c('enroll_mri', 'enroll_bld', 'enroll_np')], na.rm = T) #calculate sum
  fu_df$enroll <- ifelse(fu_df$enroll >= 2, 1, 0)
  
  #determine date of 2/3 of mri, blood, and psy completed
  enroll_dates <- c('enroll_mri_date','enroll_bld_date', 'enroll_np_date')
  fu_df$enroll_date <- NA
  for (n in 1:nrow(fu_df)){
    dates <- fu_df[n, enroll_dates] # pull out component dates
    dates <- dates[order(dates)] # order dates
      if (length(dates[!is.na(dates)]) >= 2) { # if there are 2 or more components
        fu_df$enroll_date[n] <- dates[[2]]   # pull out the second component dates
      } else {
        fu_df$enroll_date[n] <- NA         # otherwise print 'NA'
      }
    }
  fu_df$enroll_date <- as.Date(fu_df$enroll_date, origin = "1970-01-01") #turn back into date format

#determine if enrollment was met this month
  fu_df$enroll_mth <- as.character(substr(fu_df$enroll_date, 1, 7)) #store month and year, as character
  fu_df$enroll_mthT <- fu_df$enroll_mth %in% fu_df$currentYm_str #logical T or F- was criteria met this month?
  
  #Sooo I'll calculate stuff kind of this way: 
  sum(fu_df$enroll_mri[fu_df$site.x=='UP'], na.rm=TRUE)
  
  
  
  
  
#follow up recruit table (STILL NEED TO CHANGE THE VARIABLE NAMES, BUT REALLY DON'T WANT TO RIGHT THIS MOMENT)
  
# create vector of enrollment variables
enroll_vars <- c(
  "enroll_np_mthT",
  "enroll_np",
  "enroll_bld_mthT",
  "enroll_bld", 
  "enroll_mri_mthT",
  "enroll_mri"
  ) 

#turn all NAs into 0
  recruit_df[, enroll_vars] <- apply(recruit_df[, enroll_vars], 2, function(x){replace(x, is.na(x), 0)})

#initialize dataframe and names columns and rows
enroll_table <- data.frame(matrix(ncol=11, nrow=length(enroll_vars)))

#names of columns and rows on table
names(enroll_table) <- c(paste0(c('n', '%', 'n', '%', 'n', '%', 'n', '%', 'n', '%', 'total')))
row.names(enroll_table) <- c(
  "completed neuropsych current month",
  "completed neuropsych to date", 
  "completed blood current month",
  "completed blood to date",
  "completed MRI current month",
  "completed MRI to date")

# initialize counters (j = row, k = column)
j <- 1
k <- 1

for (var in enroll_vars) {
    for (site in sites) {
        enroll_table[j,k] <- sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, var], na.rm = TRUE)
        k <- k + 2
    }
  k <- 1
  j <- j + 1
}

# add in totals
j <- 1
k <- 1
for (var in enroll_vars) {
  enroll_table[j,11] <- sum(recruit_df[recruit_df$enroll == 1, var], na.rm = TRUE)
  j <- j + 1
}

#add in percentages - neuropsych current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_np_mthT'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      enroll_table[1,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

#add in percentages - neuropsych to date
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_np'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      enroll_table[2,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

#add in percentages - blood current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_bld_mthT'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      enroll_table[3,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
  }

#add in percentages - blood to date
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_bld'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      enroll_table[4,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
  }

#add in percentages - mri current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_mri_mthT'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      enroll_table[5,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
  }

#add in percentages - mri to date
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_mri'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      enroll_table[6,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

  
```


```{r recruit_table, include = FALSE, warning = FALSE}

# create vector of recruitment variables
recruit_vars <- c(
  "meta_consent_mthT.x",       #consented current month
  "consented.x",               #consented to date
  "enroll_mthT",               #enrolled current month        no
  "enroll",                    #enrolled to date              no
  "meta_complete_mthT.x",      #completed current month
  "completed.x",               #completed to date
  "terminate_baseline.x",      #terminated baseline
  "terminate_FU_06_MTH.x",     #terminated 6 mth FU
  "terminate_FU_24_MTH.x")     #terminated 24 mth FU
  
#turn all NAs into 0
  recruit_df[, recruit_vars] <- apply(recruit_df[, recruit_vars], 2, function(x){replace(x, is.na(x), 0)})
  
#initialize dataframe and names columns and rows
recruit_table <- data.frame(matrix(ncol=11, nrow=length(recruit_vars)))

#names of columns and rows on demo_table
names(recruit_table) <- c(paste0(c('n', '%', 'n', '%', 'n', '%', 'n', '%', 'n', '%', 'total')))
row.names(recruit_table) <- c(
  "consented current month",
  "consented to date", 
  "enrolled current month",
  "enrolled to date",
  "completed current month",
  "completed to date",
  "terminated during baseline",
  "terminated during 6 mth FU",
  "terminated during 24 mth FU")

# initialize counters (j = row, k = column)
j <- 1
k <- 1

for (var in recruit_vars) {
    for (site in sites) {
        recruit_table[j,k] <- sum(recruit_df[recruit_df$site.x == site, var], na.rm = TRUE)
        k <- k + 2
    }
  k <- 1
  j <- j + 1
}

# add in totals
j <- 1
k <- 1
for (var in recruit_vars) {
  recruit_table[j,11] <- sum(recruit_df[var], na.rm = TRUE)
  j <- j + 1
}

# add in percentages
# consented current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site, 'meta_consent_mthT.x'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      recruit_table[1,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

# consented total
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site, 'consented.x'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      recruit_table[2,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

# enrollment current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site, 'enroll_mthT'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      recruit_table[3,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
  }

# enrollement total
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site, 'enroll'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      recruit_table[4,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }
```

```{r recruitment_table, echo = FALSE, warning = FALSE}

#add row names to make like demo table
recruit_table <- cbind(metric = rownames(recruit_table), recruit_table)
rownames(recruit_table) <- NULL

#turns NAs into dashes
options(knitr.kable.NA = '--')

#table
kable(recruit_table, "html", escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", "condensed", full_width = T) %>%
  add_header_above(c(" ", "CU" = 2, "LA" = 2, "UP" = 2, "UT" = 2, "WU" = 2, " ")) %>%
  group_rows('Consent[note]', 1, 2, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Enrollment[note]", 3, 4, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Completion[note]", 5, 6, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Termination[note]", 7, 9, label_row_css = "background-color: #ddd; color: 999;") %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  add_footnote(c("`Consent` counts all participants who signed an OPTIMUM-Neuro consent form.", 
                 "`Enrollment` counts all participants who completed ≥2/3 of the MRI, bloodwork, and neuropsych at the baseline visit.",
                 "`Completion` counts all enrolled participants who completed the 24 mth FU visit.", 
                 "`Termination` counts participants coded as: 'not eligible for randomization', 'withdrew', 'lost to follow-up', 'death', 'investigator discretion', and 'other'. Termination does not capture participants who completed the study, i.e., this category is independent from `Completion`."), 
               notation = "number") %>%
  footnote(general = "Current month captures calendar month. The `%` columns calculate percentage of target, with all sites having common targets.")
              
rm(recruit_table)

```

------

<br>

###Enrollment breakdown

```{r enrollment_breakdown, echo = FALSE, warning = FALSE}

# create vector of enrollment variables
enroll_vars <- c(
  "enroll_np_mthT",
  "enroll_np",
  "enroll_bld_mthT",
  "enroll_bld", 
  "enroll_mri_mthT",
  "enroll_mri"
  ) 

#turn all NAs into 0
  recruit_df[, enroll_vars] <- apply(recruit_df[, enroll_vars], 2, function(x){replace(x, is.na(x), 0)})

#initialize dataframe and names columns and rows
enroll_table <- data.frame(matrix(ncol=11, nrow=length(enroll_vars)))

#names of columns and rows on table
names(enroll_table) <- c(paste0(c('n', '%', 'n', '%', 'n', '%', 'n', '%', 'n', '%', 'total')))
row.names(enroll_table) <- c(
  "completed neuropsych current month",
  "completed neuropsych to date", 
  "completed blood current month",
  "completed blood to date",
  "completed MRI current month",
  "completed MRI to date")

# initialize counters (j = row, k = column)
j <- 1
k <- 1

for (var in enroll_vars) {
    for (site in sites) {
        enroll_table[j,k] <- sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, var], na.rm = TRUE)
        k <- k + 2
    }
  k <- 1
  j <- j + 1
}

# add in totals
j <- 1
k <- 1
for (var in enroll_vars) {
  enroll_table[j,11] <- sum(recruit_df[recruit_df$enroll == 1, var], na.rm = TRUE)
  j <- j + 1
}

#add in percentages - neuropsych current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_np_mthT'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      enroll_table[1,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

#add in percentages - neuropsych to date
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_np'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      enroll_table[2,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

#add in percentages - blood current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_bld_mthT'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      enroll_table[3,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
  }

#add in percentages - blood to date
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_bld'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      enroll_table[4,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
  }

#add in percentages - mri current month
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_mri_mthT'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'per_month']
      enroll_table[5,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
  }

#add in percentages - mri to date
j <- 2
  for (site in sites) {
      n = sum(recruit_df[recruit_df$site.x == site & recruit_df$enroll == 1, 'enroll_mri'])
      d = targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'target']
      enroll_table[6,j] <- sprintf("%1.0f%%", 100*round(n/d, 2))
      j <- j + 2
    }

```

```{r enrollment_table, echo = FALSE, warning = FALSE}

#add row names to make like demo table
enroll_table <- cbind(metric = rownames(enroll_table), enroll_table)
rownames(enroll_table) <- NULL

#turns NAs into dashes
options(knitr.kable.NA = '--')

#table
kable(enroll_table, "html", escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", "condensed", full_width = T) %>%
  add_header_above(c(" ", "CU" = 2, "LA" = 2, "UP" = 2, "UT" = 2, "WU" = 2, " ")) %>%
  group_rows('Neuropsych[note]', 1, 2, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Blood[note]", 3, 4, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("MRI[note]", 5, 6, label_row_css = "background-color: #ddd; color: 999;") %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  add_footnote(c(
     "`Neuropsych` counts all participants who completed both the RBANS and D-KEFS in full.",
     "`Blood` counts all participants who submitted one aliquot of plasma.",
     "`MRI` counts all participants who completed the T1."), 
               notation = "number") %>%
  footnote(general = "Current month captures calendar month. The `%` columns calculate percentage of target, with all sites having common targets.")
              
rm(enroll_table)

```

------

<br>

###Demographics 

```{r demographics, echo = FALSE, warning = FALSE}

# create a separate dataframe of data from just baseline timepoint
demo_df <- df[(df$timepoint == 'baseline'),]

# modify race variable to have 2 levels
demo_df$demo_race <- ifelse(demo_df$demo_race != 2, 1, 2)

# create vector of demographic variables
demo_vars <- c(
  'demo_sex',
  'demo_ethnicity',
  'demo_race',
  'demo_age',
  'demo_edu')

# initialize dataframe 
demo_table <- data.frame(matrix(ncol=6, nrow=length(demo_vars)))

# names of columns and rows on demo_table
names(demo_table) <- c(paste0(c('CU', 'LA', 'UP', 'UT', 'WU', 'p')))
row.names(demo_table) <- c('sex (M:F:O)', 'ethnicity (H:NH)', 'race (minority:non-minority)', 'age', 'education')

# initialize counters (j = row, k = column)
j <- 1
k <- 1

# for loop for categorical
    for (var in demo_vars[1:3]) {
      
      # count observations for each cluster
          for (site in sites) {
            out <- table(demo_df$site, demo_df[[var]])
            demo_table[j,k] <- paste(out[site,], collapse = ':')
            k <- k + 1}
      
     # run chi-square tests 
          chi_sq <- chisq.test(out)
          demo_table[j,k] <- sub("^(-?)0.", "\\1.", sprintf("%.3f", chi_sq$p.value))
          
          k <- 1
          j <- j + 1} 

# initialize counters (j = row, k = column)
j <- 4
k <- 1

# for loop for continuous variables
    for (var in demo_vars[4:5]) {
      
      # calculate means and SDs for each site
          for (site in sites) {
            M <- sprintf('%.02f', mean(demo_df[demo_df$site == site, var], na.rm = TRUE))
            SD <- sprintf('%.02f', sd(demo_df[demo_df$site == site, var], na.rm = TRUE))
            demo_table[j,k] <- paste( M,' (',SD,')', sep='') 
            
            k <- k + 1}
      
      # run one-way ANOVA with cluster as between-subjects variable
          F_test <- aov(demo_df[[var]] ~ demo_df$site, na.action=na.omit)
      # extract p-value
          F_test.p.value <- summary(F_test)[[1]][["Pr(>F)"]][[1]]
      # rounded p-value to 3 decimals and without leading zero
          F_test.p <- sub("^(-?)0.", "\\1.", sprintf("%.3f", F_test.p.value))
          demo_table[j,k] <- F_test.p 
          
          k <- 1 
          j <- j + 1}

```

```{r demographics_table, echo = FALSE, warning = FALSE}

demo_table$p <- as.numeric(demo_table$p)

#table
demo_table %>%
  mutate(
    metric = row.names(demo_table),
    p = cell_spec(p, "html", color = ifelse(p < 0.05, "red", "black"))
    ) %>%
  select(metric, CU, LA, UP, UT, WU, p) %>%
  kable('html', escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", full_width = T) %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  footnote(general = "These values represent all participants consented at the OPTIMUM-Neuro baseline visit. It does not take into account the participants who consented but did not meet enrollment criteria, and/or who later terminated. Values in brackets are standard deviation. The _p_ values result from Chi-squared for categorical variables and omnibus ANOVA for continuous variables.")

rm(demo_df, demo_table)

```


