---
title: "OPTIMUM-Neuro conference call 2019-02-15"
output: html_document
always_allow_html: yes
---

```{r setup, include = FALSE, warning = FALSE}

# load graphics libraries
    library(plyr)
    library(ggplot2)
    library(reshape2)
    library(tidyverse) 
    library(knitr)
    library(kableExtra)
    library(plotly)
    library(stringr)
    library(bsselectR)
    library(lubridate)
    library(textclean)

   # setwd('/projects/gherman/OPT_callReport')

```

```{r data cleaning, include = FALSE, warning = FALSE}

source('functions.R')

#read in data
  df <- read.csv('/home/gherman/Downloads/OPTIMUMMainDatabaseF_DATA_2019-06-19_1022.csv', na.strings=c("","NA")) #contains all but MRI
  mri <- read.csv('/home/gherman/Downloads/OPTIMUMMainDatabaseF_DATA_2019-06-19_1023.csv', na.strings=c("","NA")) #contains just MRI
  targets <- read.csv('info/OPT_recruitTargets.csv')
  #make sure target data is in proper format
  targets$month <- as.character(targets$month)
  
    sites <- factor(x = c('CU', 'LA', 'UP', 'UT', 'WU'), levels = c('CU', 'LA', 'UP', 'UT', 'WU'))
  

  
  df <- prepare_init_df(df, mri, sites)
  
```
>We are currently in month `r targets[targets$month == as.character(substr(Sys.Date(), 1, 7)), 'month_count']` of the study.  

This report is meant to provide a quick overview of the OPTIMUM-Neuro study, to facilitate discussion on its bi-weekly conference calls. All data summarized here derives from the 'tracking' arm of the WashU REDCap database (arm 6), and thus is only as accurate as the data captured there.
<br> 

------
<br> 

###Recruitment

```{r recruitment, echo = FALSE, include = FALSE, warning = FALSE}

#make a smaller dataframe
recruit_df <- df[, c(
  "record_id", 
  "site",
  "timepoint",
  
  "meta_consent_date", 
  "meta_terminate_date",
  "meta_terminate_reason",
  
  "plasma_blood_status", 
  "plasma_blood_date",
  
  "mr_t1",
  "mr_date",
  
  "dkefs_complete",
  "dkefs_date_admin",
  "rbans_complete",
  "rbans_date_admin")] 

recruit_df <- prepare_recruit_df(recruit_df)
 
 
```
 
```{r enrolment, echo = FALSE, include = FALSE, warning = FALSE}
#add column that indicates if enrolled (where enrollment is 2 of 3 mri, blood, neuropsych)
  enroll_df <- prepare_enroll_df(recruit_df) #make a new dataframe (did this because want new DF to consider only baseline)

 
 
#merge back dataframes #what's the purpose of merging it here?
recruit_df <- merge(recruit_df, enroll_df, by = c('record_id', 'timepoint'), all.x = TRUE)
  
```

  

```{r recruit_table, warning = FALSE, echo = FALSE}

recruit_table <- make_recruit_table(recruit_df, sites, targets)

#turns NAs into dashes
options(knitr.kable.NA = '--')

#table
kable(recruit_table, "html", escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", "condensed", full_width = T) %>%
  add_header_above(c(" ", "CU" = 2, "LA" = 2, "UP" = 2, "UT" = 2, "WU" = 2, " ")) %>%
  group_rows('Consent[note]', 1, 2, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Enrollment[note]", 3, 4, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Completion[note]", 5, 6, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Termination[note]", 7, 9, label_row_css = "background-color: #ddd; color: 999;") %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  add_footnote(c("`Consent` counts all participants who signed an OPTIMUM-Neuro consent form.", 
                 "`Enrollment` counts all participants who completed ≥2/3 of the MRI, bloodwork, and neuropsych at the baseline visit.",
                 "`Completion` counts all enrolled participants who completed the 24 mth FU visit.", 
                 "`Termination` counts participants coded as: 'not eligible for randomization', 'withdrew', 'lost to follow-up', 'death', 'investigator discretion', and 'other'. Termination does not capture participants who completed the study, i.e., this category is independent from `Completion`."), 
               notation = "number") %>%
  footnote(general = "Current month captures calendar month. The `%` columns calculate percentage of target, with all sites having common targets.")

```

------

<br>

###Enrollment breakdown

```{r enrollment_table, echo = FALSE, warning = FALSE}

  enroll_table <- make_enroll_table(recruit_df, sites, targets)

#add row names to make like demo table
enroll_table <- cbind(metric = rownames(enroll_table), enroll_table)
rownames(enroll_table) <- NULL

#turns NAs into dashes
options(knitr.kable.NA = '--')

#table
kable(enroll_table, "html", escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", "condensed", full_width = T) %>%
  add_header_above(c(" ", "CU" = 2, "LA" = 2, "UP" = 2, "UT" = 2, "WU" = 2, " ")) %>%
  group_rows('Neuropsych[note]', 1, 2, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("Blood[note]", 3, 4, label_row_css = "background-color: #ddd; color: 999;") %>%
  group_rows("MRI[note]", 5, 6, label_row_css = "background-color: #ddd; color: 999;") %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  add_footnote(c(
     "`Neuropsych` counts all participants who completed both the RBANS and D-KEFS in full.",
     "`Blood` counts all participants who submitted one aliquot of plasma.",
     "`MRI` counts all participants who completed the T1."), 
               notation = "number") %>%
  footnote(general = "Current month captures calendar month. The `%` columns calculate percentage of target, with all sites having common targets.")
              
```

------

<br>

###Demographics 

```{r demographics, echo = FALSE, warning = FALSE}

demo_table <- make_demo_table(df, sites)

```

```{r demographics_table, echo = FALSE, warning = FALSE}

demo_table$p <- as.numeric(demo_table$p)

#table
demo_table %>%
  mutate(
    metric = row.names(demo_table),
    p = cell_spec(p, "html", color = ifelse(p < 0.05, "red", "black"))
    ) %>%
  select(metric, CU, LA, UP, UT, WU, p) %>%
  kable('html', escape = F, align = c('l', 'c', 'c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", full_width = T) %>%
  column_spec(1, width = "20em", background = '#f5f5f5') %>%
  footnote(general = "These values represent all participants consented at the OPTIMUM-Neuro baseline visit. It does not take into account the participants who consented but did not meet enrollment criteria, and/or who later terminated. Values in brackets are standard deviation. The _p_ values result from Chi-squared for categorical variables and omnibus ANOVA for continuous variables.")

rm(demo_df, demo_table)

```




```{r}


#note to self.... the enrollmrimth thing doesnt really work as the only indicator, it should really be combined with mr_t1 (which is enroll_mri i guess) to give an accurate picture of numbers. 


fu_due_list <- recruit_df%>% select(record_id,mri_fu_due, np_fu_due, bld_fu_due) %>% filter(!is.na(mri_fu_due) | !is.na( np_fu_due) | !is.na( bld_fu_due))

fu_df <- recruit_df %>% filter(.,timepoint=="6 mth FU") %>% select(.,-c(enroll_mri, enroll_mri_date)) 

fu_df <- prep_fu_df(fu_df) %>% merge(.,fu_due_list, by=c('record_id'))

  #Sooo I'll calculate stuff kind of this way: 
  sum(fu_df$enroll_mri[fu_df$site.x=='UP'], na.rm=TRUE)
  
```


I guess I can remove the percentages here, although I may want to do percentage of people who consented 8 or more months ago who are obtaining follow-up!! That'd be cool.

In order to do that, I'll need a df with participant IDs, initial scan date, later scan date, whether it's >= 6 months earlier. 

AFTER HERE IS THE NOT REAL STUFF THAT I SHOULD DELETE OR MOVE AROUND LATER
```{r}

fu_table <- make_fu_table(fu_df, sites,targets)

```




```{r}
df1 <- fu_df %>% select(record_id, fu_mri = enroll_mri, fu_mri_date = enroll_mri_date)


df2 <- enroll_df %>% select(record_id, enroll_mri, enroll_mri_date)

#testmerged <- merge(df2 ,df1,by=c('record_id'), all=TRUE) %>%
#  mutate(., fu_due = ifelse(enroll_mri_date %m+% months(6) < Sys.Date(), 1, 0))

#fu_due_list <- testmerged %>% select(record_id,fu_due) 

#fu_df2 <- fu_df %>% merge(.,fu_due_list, by=c('record_id'))


```



TODO: change the MRI coding (because we have options 1 2 and 3 now in the neuro mri form), and also make it not rely on mri date




